Fast Sorting on Modern Computers

### 2005-08-31 docs/tech_report/conclusions.tex
davids fixes to heap sort due to incorrect taken vs not-taken. also, first corrections to conclusions

\label{conclusions}
This section presents a summary of the results from previous chapters, compares
our results with LaMarca's, lists important results and discoveries, and
suggests contributions made to computer science by this research.

@@

We found that insertion sort has the best performance of any of the elementary
sorts, with fewer instructions and cache misses than any other elementary sort,
and exactly one branch misprediction per key. Selection sort also has very few
branch mispredictions, and our results match figures created using Knuth's
harmonics formula. We found that bubblesort and shakersort have poor branch
prediction performance, and that this cripples their performance, especially
compared to selection sort which is otherwise similar in behaviour. We
determined that bubblesort has better branch performance at the start and end of
its sorting.

We also found that reducing the number of runs performed by shakersort and
bubblesort is counter productive due to additional book-keeping cost. We found
also that of the elementary sorts, only insertion sort has good cache
performance, as the others do not exploit any temporal locality.

@@

Additionally, we created an improvement to heapsort using a 2-heap, which
removed the need for a sentinel, without introducing any more instructions. We
found that a $k$-ary heap has similar branch prediction performance to selection
sort, though not exactly, in that the fanout becomes more predictable the larger
it is.

@@

We found that algorithm S has a lower instruction count than algorithm N, but
that this does not necessarily translate to better performance. Our results show
also that the level 1 cache needs to be taken into account, that the extra
instruction in multi-mergesort make it slower than tiled mergesort, and that
multi-mergesort has regular access patterns during the k-way merge, which can be
taken advantage of by a two-level adaptive predictor.

@@

Our results show base quicksort being the fastest of the quicksorts. We note our
lowest instruction count occurs with the median-of-three partitioning. We also
see that using a binary search slows multi-quicksort down compared to a
sequential search due to branch mispredictions, but that its extra instructions
make it slower than memory-tuned quicksort. Despite being slower in our tests,
our simulations show memory-tuned quicksort to have fewer cache misses,
instructions and branch mispredictions than base quicksort.

>>>>

\label{conclusions}
This section presents a summary of the results from previous chapters, lists
important branch prediction results and discoveries, compares our results with
LaMarca's and discusses important cache results, discusses the best sort that we
found in a number of catagories, and suggests contributions made to computer
science by this research.

@@

We found that insertion sort has the best performance of any of the elementary
sorts, with fewer instructions and cache misses than any other elementary sort,
and exactly one branch misprediction per key. Selection sort also has very few
branch mispredictions, and our results match figures created using Knuth's
harmonics formula. We found that bubblesort and shakersort have poor branch
prediction performance, and that this cripples their performance, especially
compared to selection sort which is otherwise similar in behaviour. We
determined that bubblesort has better branch performance at the start and end of
its sorting, then in the middle.

We also found that reducing the number of runs performed by shakersort and
bubblesort is counter-productive due to additional book-keeping cost. We found
that of the elementary sorts, only insertion sort has good cache performance, as
the others do not exploit any temporal locality.

@@

Additionally, we created an improvement to heapsort using a 2-heap, which
removed the need for a sentinel, without introducing any more instructions. We
found that a $k$-ary heap has similar, though not the same, branch prediction
performance as selection sort, in that the fanout becomes more predictable the
larger it is.

@@

We found that algorithm S has a lower instruction count than algorithm N, but
that this does not translate to better performance. Our results show also that
the level 1 cache needs to be taken into account, both in tiling and in
alignment; that the extra instructions in multi-mergesort make it slower than
tiled mergesort, and that multi-mergesort has regular access patterns during the
k-way merge, which can be taken advantage of by a two-level adaptive predictor.

@@

Our results show base quicksort being the fastest of the quicksorts. We note our
lowest instruction count occurs with median-of-three partitioning. We also see
that using a binary search slows multi-quicksort down compared to a sequential
search due to branch mispredictions, but that both have too many extra
instructions to compete against memory-tuned quicksort. Despite being slower in
our tests, our simulations show memory-tuned quicksort to have fewer cache
misses, instructions and branch mispredictions than base quicksort.
